Forecasting Gamma-Ray Bursts using Gravitational Waves

### 2018-04-15 Paper/Manuscript.tex
Finished Conor's corrections, added x-axes label to Fig. 5 and one more reference.

\author{Sarp Akcay}
\affiliation{School of Mathematics \& Statistics, University College Dublin, Belfield, Dublin 4, Ireland.}

\date{\today}

@@

we obtain $e\approx 6\times 10^{-5}$. Thus our setting $e=0$ is very reasonable.

Let us briefly discuss the possibility of eccentric BNS systems {\bf FINISH}
\cite{1975MNRAS.172P..15F,Grindlay:2005ym}

@@

\section{GW170817 as a test of our model for the Advanced LIGO-Virgo network}\label{sec:GW170817}
On 17 August 2017, at 12:41:04 UTC, GWs from the inspiral and merger of a binary neutron star system %were detected by the Advanced LIGO-Virgo network.
%The event, first picked up by LIGO Hanford (H1), 
swept the ALV network's frequency band from $\approx 30\,$Hz to $\approx 2000\,$Hz accumulating a total SNR of
%18.8, 26.4 and 2.0 at H1, LIGO Livingstone (L1) and Virgo (V1), respectively, yielding 
$\rho_\text{tot} = 32.4$ \cite{GW170817}
The inferred duration of the event in the network bandwidth was $\approx 57\,$ seconds during which $\approx 3000$ cycles of GW were emitted \cite{GW170817_Facts}.
Assuming low-spin priors  ($\chi \le 0.05$, see point 7 of Sec.~\ref{sec:idealizations}), the following values were inferred for the total mass, the chirp mass,
and the luminosity distance of the source with 90\% confidence:
%
\be
M = 2.74^{+0.04}_{-0.01} M_\odot,\quad M_c = 1.188^{+0.004}_{-0.002} M_\odot,\quad D= 40^{+8}_{-14}\,\text{Mpc}. \label{eq:GW170817_params}
\ee
%
Note that the error bar for the chirp mass is very small
due to the fact that it can be extracted to high precision over thousands of GW cycles via Eq.~(\ref{eq:fdot}).

@@

Fig.~\ref{fig:figGW170817} shows our network. The red curve is L1's ASD (sensitivity) during GW170817. We assume our three IFOs to identically have this sensitivity,

@@

 as A-LIGO horizon will be slightly over 200\,Mpc and A-Virgo's around 130\,Mpc.

@@

Given that we might at best have a couple of minutes of warning time for a once-a-decade event, we do not expect to be able forecast GRBs in the Advanced LIGO-Virgo era. 

@@

Although the results from each numerical study of BH-NS mergers may somewhat vary what is clear is that we must have high spin and low mass for the black hole for tidal disruption to happen. 
We are thus led to set $M_\text{BH} =5 M_\odot$. 
This raises an interesting astrophysical question as to whether it is possible to have such light-weight high-spin black holes assuming that
they are spun up by accretion from their secondary (that will later on become a NS) in binary systems, thus they will have to gain mass.
It is possible that the mass gain pushes $M_\text{BH}$ too high to either cause a direct plunge of the NS before TD or decrease $\tau_\text{insp}$ considerably
to eliminate any chances of getting an advance warning.
However, if the spin of the BH comes from the supernova explosion mechanism itself as argued for G1915 in Ref.~\cite{McClintock:2006xd} then it is possible to have a 
low-mass BH in a binary system\footnote{We thank Dr. Morgan Fraser
for pointing this out.}.

@@

\acknowledgements 
SA thanks Conor O'Toole for endless feedback and Niels Warburton for a careful reading of this article.

>>>>

\author{Sarp Akcay}
\affiliation{School of Mathematics \& Statistics, University College Dublin, Belfield, Dublin 4, Ireland.}
\affiliation{Friedrich-Schiller-Universit{\"a}t Jena}
%\date{\today}

@@

we obtain $e\approx 6\times 10^{-5}$. Thus our setting $e=0$ is very reasonable\footnote{
It is possible to have 
high-eccentricity binary black hole mergers in the vicinity of galactic nuclei due close encounter captures \cite{OLeary:2008myb}.
However, it is not clear how much this applies to BNS systems.}.
%for binary systems
%non-zero eccentricity binary mergers
%for sytems that have formed by tidal capture mechanisms in globular clusters {\bf FINISH}
%\cite{1975MNRAS172P15F,Grindlay:2005ym}
%but these binaries are expected to be outnumbered significantly by the circularized ones.

@@

\section{GW170817 as a test of our model for the Advanced LIGO-Virgo network}\label{sec:GW170817}
On 17 August 2017, at 12:41:04 UTC, GWs from the inspiral and merger of a binary neutron star system %were detected by the Advanced LIGO-Virgo network.
%The event, first picked up by LIGO Hanford (H1), 
swept the ALV network's frequency band from $\approx 30\,$Hz to $\approx 2000\,$Hz, accumulating a total SNR of
%18.8, 26.4 and 2.0 at H1, LIGO Livingstone (L1) and Virgo (V1), respectively, yielding 
$\rho_\text{tot} = 32.4$ \cite{GW170817}.
The inferred duration of the event in the network bandwidth was $\approx 57\,$ seconds during which $\approx 3000$ cycles of GWs were emitted \cite{GW170817_Facts}.
Assuming low-spin priors  ($\chi \le 0.05$, see point 7 of Sec.~\ref{sec:idealizations}), the following values were inferred for the total mass, the chirp mass,
and the luminosity distance of the source with 90\% confidence:
%
\be
M = 2.74^{+0.04}_{-0.01} M_\odot,\quad M_c = 1.188^{+0.004}_{-0.002} M_\odot,\quad D= 40^{+8}_{-14}\,\text{Mpc}. \label{eq:GW170817_params}
\ee
%
Note that the error bar for the chirp mass is very small
due to the fact that it can be extracted to high precision over thousands of GW cycles via Eq.~(\ref{eq:fdot}).

@@

Fig.~\ref{fig:figGW170817} shows our network. The red curve is L1's ASD (sensitivity) during GW170817. We assume our three IFOs have identical sensitivities,

@@

 as A-LIGO's horizon will be slightly over 200\,Mpc and A-Virgo's around 130\,Mpc.

@@

Given that we might at best have a couple of minutes of warning time for a potentially once-a-decade event, we do not expect to be able forecast GRBs in the Advanced LIGO-Virgo era. 

@@

Although the results from each numerical study of BH-NS mergers may somewhat vary what is clear is that we must have high spin and low mass for the black hole for tidal disruption to happen. 
We are thus led to set $M_\text{BH} =5 M_\odot$. 
This raises an interesting astrophysical question as to whether it is possible to have such light-weight high-spin black holes assuming that
they are spun up by accretion from their secondary (that will later on become a NS) in binary systems, thus they will have to gain mass.
It is possible that the mass gain pushes $M_\text{BH}$ too high to either cause a direct plunge of the NS before TD or decrease $\tau_\text{insp}$ considerably
to eliminate any chances of getting an advance warning.
However, if the spin of the BH comes from the supernova explosion mechanism itself as argued for G1915 in Ref.~\cite{McClintock:2006xd} then it is possible to have a 
low-mass BH in a binary system\footnote{We thank Dr. Morgan Fraser of UCD Astrophysics
for pointing this out.}.

@@

\acknowledgements 
SA thanks Conor O'Toole and Niels Warburton for endless feedback and a thorough perusal of this article.
